Supplementary Information for 'Detecting' Manuscript

### 2023-11-14 SI_draft.Rmd
SI add bin widths, top panel fig 2, update captions, number of replicates

```{=tex}
\beginsupplement
```
# $\lambda$ and relationship estimates

@@

![Figure S1. Absolute deviation in estimate from known value of $\lambda$. Points are jittered horizontally for visualization, and are deviations for each simulated replicate (n = 1000) across varying sample sizes of body size values. Lines are the average absolute deviation from the estimate. Colors indicate which method was used to estimate $\lambda$.](figures/abs_dev_by_body_n.png)
```{r, echo=FALSE, results='asis', message=FALSE}
library(knitr)
table1 <- read_csv("tables/abs_dev_n_vary.csv") %>%
  mutate(diff_abs_mean = round(diff_abs_mean, 3),
         diff_abs_sd = round(diff_abs_sd, 3)) %>%
  rename(method = name,
         "Body sample size" = body_n,
         "mean abs diff"=diff_abs_mean,
         "SD abs diff"=diff_abs_sd)
kable(table1,
         caption = "Table S1. Deviation of estimates across methods for different sample sizes of body size values. The mean and standard deviation for each row (mean abs diff and sd abs diff, respectively) in the table are calculated across all replicates for each value of lambda (1000 reps * 5 lambda values = 5000 values.",
      padding = 10)
```

@@

```{r, echo=FALSE, results='asis', message=FALSE}
library(knitr)
table2 <- read_csv("tables/si_table2.csv") %>%
  mutate(diff_abs_mean = round(diff_abs_mean, 3),
         diff_abs_sd = round(diff_abs_sd, 3)) %>%
  rename(method = name,
         "mean abs diff"=diff_abs_mean,
         "SD abs diff"=diff_abs_sd)
kable(table2,
         caption = "Table S2. Deviation of estimates across methods when the values of the hypothetical gradient (large x) and the range of body sizes (small m) are changed.")
```

@@

>>>>

```{=tex}
\beginsupplement
```

# ELBn bin widths

The ELBn method uses six bins of equal width on a logarithmic scale. To illustrate how the widths of these bins varies, we provide the following example when the body sizes range from 1 to 100 g vs 1 to 100g.

```{r, echo = FALSE, message=FALSE}
dw_range_100 <- c(1, 100)

elbn_100 <- exp(seq(floor(log(min(dw_range_100))),
                          ceiling(log(max(dw_range_100))),
                          length.out = 7))

dw_range_1000 <- c(1, 1000)

elbn_1000<- exp(seq(floor(log(min(dw_range_1000))),
                          ceiling(log(max(dw_range_1000))),
                          length.out = 7))

elb_df <- data.frame(x = c(elbn_100[1:6], elbn_1000[1:6]),
           xend = c(elbn_100[2:7], elbn_1000[2:7]),
          y  = rep(c("range = 100", "range = 1000"), each= 6),
          yend = rep(c("range = 100", "range = 1000"), each= 6),
          color = rep(c("black", "red"), 6))

l2n_100 <- 2^(floor(log2(min(dw_range_100))):
                  ceiling(log2(max(dw_range_100))) )

l2n_1000 <- 2^(floor(log2(min(dw_range_1000))):
                  ceiling(log2(max(dw_range_1000))) )

# length(l2n_100)
# length(l2n_1000)

l2n_df <- data.frame(x = c(l2n_100[1:7], l2n_1000[1:10]),
           xend = c(l2n_100[2:8], l2n_1000[2:11]),
          y  = c(rep("range = 100", 7),rep("range = 1000", 10)),
          yend = c(rep("range = 100", 7),rep("range = 1000", 10)),
          color = c(rep(c("black", "red"), length.out = 7), rep(c("black", "red"), length.out = 10))) 
```

```{r, echo=FALSE}
elb_df %>%
  slice(1:5, 7:10) %>%
  ggplot(aes(x = x,
             xend = xend,
             y = y,
             yend = yend,
             color = color)) +
  geom_segment(linewidth = 3,
               linejoin = 'bevel',
               lineend = 'butt') +
  geom_point(shape = "|", size = 7, color = "black") +
  theme_bw() +
  coord_cartesian(xlim = (c(0, 50))) +
  labs(x = "body size",
       y = "") +
  theme(legend.position = "none")

```
Figure S1. The edges of the first 4 bins are shown when the body sizes range from 0-100 (bottom line) and 0-1000 (top line). The bins have alternating colors, and black segments are used to emphasize bin edges. As you can see, the range of the smallest bins are similar, but the edges quickly diverge in the large bin sizes. Only the first 4 bins are shown, and the image is cropped for visualization (i.e., the edges of the last bins are not shown).

# L2n bin widths

To contrast the variable bin widths, we show that the widths of the bins in the L2n method are always the same, but the number of bins changes to accomodate the range of body sizes present.

```{r, echo=FALSE}
l2n_df %>%
  ggplot(aes(x = x,
             xend = xend,
             y = y,
             yend = yend,
             color = color)) +
  geom_segment(linewidth = 3,
               linejoin = 'bevel',
               lineend = 'butt') +
  geom_point(shape = "|", size = 7, color = "black") +
  theme_bw() +
  #coord_cartesian(xlim = (c(0, 50))) +
  labs(x = "body size",
       y = "") +
  theme(legend.position = "none") +
  scale_x_log10()
```

Figure S2. The edges of all bins for the L2n method are shown when body sizes range from 0-100 (bottom line) and 0-1000 (top line). The bins have alternating colors, and black segments are used to emphasize bin edges. As you can see, the bin widths are identical, but the number of bins increases with the range of body sizes in the data. The plot is presented on a log~10~ scale for visualization.

# Top panel of Figure 2, main text

![](figures/SI_fig2_top_panel.png)
Figure S3. The variation in the distributions of $\lambda$ estimates for "steep" values in Figure 2 in the main text makes it difficult to compare them with the "shallower" distirbutions. Here, we repeat just the top panel of Figure 2 from the main text to show the variation in estimates across methods. As in the other values of $\lambda$, MLE is the most accurate and precise. The L2N method performs slightly worse, and the ELBn method is much more variable, likely due to the fact that the presence/absence of even a single large individual could drastically alter the widths of the bins present within an individual simulation replicate (See main text).

# $\lambda$ and relationship estimates

@@

![Figure S4. Absolute deviation in estimate from known value of $\lambda$. Points are jittered horizontally for visualization, and are deviations for each simulated replicate (n = 1000) across varying sample sizes of body size values. Lines are the average absolute deviation from the estimate. Colors indicate which method was used to estimate $\lambda$.](figures/abs_dev_by_body_n.png)

```{r, echo=FALSE, results='asis', message=FALSE}
library(knitr)
table1 <- read_csv("tables/abs_dev_n_vary.csv") %>%
  mutate(diff_abs_mean = round(diff_abs_mean, 3),
         diff_abs_sd = round(diff_abs_sd, 3)) %>%
  rename(method = name,
         "Body sample size" = body_n,
         "mean abs diff"=diff_abs_mean,
         "SD abs diff"=diff_abs_sd)
kable(table1,
         caption = "Table S1. Deviation of estimates across methods for different sample sizes of body size values. The mean and standard deviation for each row (mean abs diff and sd abs diff, respectively) in the table are calculated across all replicates for each value of lambda (1000 reps * 5 lambda values = 5000 values.",
      padding = 10)
```

@@

```{r, echo=FALSE, results='asis', message=FALSE}
library(knitr)
table2 <- read_csv("tables/si_table2.csv") %>%
  mutate(diff_abs_mean = round(diff_abs_mean, 3),
         diff_abs_sd = round(diff_abs_sd, 3)) %>%
  rename(method = name,
         "mean abs diff"=diff_abs_mean,
         "SD abs diff"=diff_abs_sd)
kable(table2,
         caption = "Table S2. Deviation of estimates across methods when the values of the hypothetical gradient (large x) and the range of body sizes (small m) are changed compared with the results presented in the manuscript (main). ")
```

@@

# Altering the number of replicates

The results presented in the main text for experiment 2.1 use 1000 replicates for each scenario. This number was chosen arbitrarily - we wanted a large enough number to capture the variability, but a small enough number to make repeated simulations tractable. Here, we alter the number of replicates to show how the results presented in the main text were largely invariant to this decision.

![Figure S5. We varied the number of replicates (n = 10, 50, 100, 200, 250, 500, 750, 1000) and plotted the distribution of $\beta$ estimates across the three $\lambda$ scenarios. The point represents the median of the distribution, and the bold and narrow bars represent the 66th and 95th percentile of the distributions, respectively. There was some variation in estimates when n < ~250, but generally the results were surprisingly robust to the number of replicates chosen. Note that the y-axis varies across rows.](figures/SI_vary_rep_n.png)

# Confidence intervals

We examined the proportion of 95% Confidence Intervals which had the true value for site specific $\lambda$'s (S6) as well as the CI's for the $\beta_{env}$ across $\lambda$ scenarios (S7) as well as when the effect size of $\beta_{env}$ varies.

![Figure S6. CI's and $\lambda$ estimates. CI's are colored if they contain the true value (blue) or if they do not (red). Only the first 500 replicates are displayed for visualization purposes. They are ordered by the rank of the lower 95% CI bound](figures/lamba_ci_true.png)

![Figure S7. CI's and $\beta$ estimates when varying the known value of $\beta$. CI's are colored if they contain the true value (blue) or if they do not (red). Only the first 500 replicates are displayed for visualization purposes. They are ordered by the rank of the lower 95% CI bound](figures/rel_data_ci_true.png)

![Figure S8. CI's and $\beta$ estimates when varying the scenario of $\lambda$ parameter space. CI's are colored if they contain the true value (blue) or if they do not (red). Only the first 500 replicates are displayed for visualization purposes. They are ordered by the rank of the lower 95% CI bound](figures/angles_ci_true.png)
